---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![picMaps status badge](https://pifsc-protected-species-division.r-universe.dev/badges/picMaps)](https://pifsc-protected-species-division.r-universe.dev/picMaps)

# Installation

## R-Universe

#### Binary

```{r, eval=FALSE}
install.packages('picMaps', 
                 repos=c('https://pifsc-protected-species-division.r-universe.dev','https://cloud.r-project.org')
)
```

#### Source

```{r, eval=FALSE}
install.packages('picMaps', type='source', 
                 repos=c('https://pifsc-protected-species-division.r-universe.dev','https://cloud.r-project.org')
)
```

## Github

*You will need a C++ compiler for R*

```{r, eval=FALSE}
remotes::install_github('pifsc-protected-species-division/picMaps')
```

# Initial Use

#### Data storage

To use this package you will have to first specify a storage directory for the downloaded data. This is accomplished in the following way

```{r eval=FALSE}
library(picMaps)
set_data_storage(path = "~/.picmaps_data")
```

The default directory created is `~/.picmaps_data`, but you are free to specify whatever directory you want. If it does not exist it will be created.

#### Data download

After the package is aware of the desired storage directory, you can download OSM coastline polygon data from <https://osmdata.openstreetmap.de/data/land-polygons.html> or ETOPO bathymetry data from <https://www.ncei.noaa.gov/products/etopo-global-relief-model>.

To download OSM coastline polygons run the following

```{r, eval=FALSE}
library(picMaps)
osm_download(force = FALSE, clean_shp=TRUE)
```

If `force=TRUE` the data will be download even if `picMaps` detects the data in the storage directory. When data is downloaded. The `.shp` files will be imported to `R` and transformed to a searchable `.gpkg` file. If `clean_shp=TRUE` the downloaded `.shp` files will be deleted.

To download the ETOPO bathymetry rasters, use the following

```{r, eval=FALSE}
library(picMaps)
etopo_download(resolution = 60, force = FALSE)
```

The default resolution is 60 arcsecs, but 30 arcsecs can also downloaded. Again, if the data are already present in the storage directory you must set `force=TRUE` to re-download it.

# Regular Usage

Here is an example of data usage after download. We will use the data to make a map near the main Hawaiian Islands. First we define a spatial object, `x`, for we we would like the coastline polygons. All coastline polygons that are at least partly within the bounding box will be selected from the `.gpkg` file. The argument `keep=0.2` will keep 20% of the polygon points. For large scale maps this is generally adequate.`union=FALSE` will maintain separate single polygons for land. 
```{r}
library(sf)
library(terra)
library(crsuggest)
library(picMaps)

x <- st_bbox(c(xmin=198.5, xmax=206, ymin=18.5, ymax=23), crs=4326) |>
  st_as_sfc()
prj <- as.numeric(suggest_crs(x)$crs_code[[1]])
x <- st_transform(x, prj) |> st_as_sf()

mhi <- osm_coast(x, keep=0.2, union=FALSE)
```

Now we will obtain the associated bathysphere data
```{r}
bathy <- etopo_rast(x)
bathy[bathy>0] <- NA
```

And finally, use them in a plot.
```{r ex_map}
library(ggplot2)
library(ggspatial)

tropic_ocean <- rev(c("#8aeed5","#63b9db","#4479e1","#2645e0","#071eed"))
        
ggplot() +
  layer_spatial(data=bathy, ) +
  scale_fill_gradientn(colours=tropic_ocean, name="Depth (m)", na.value = NA) +
  layer_spatial(data=mhi, fill="tan") +
  theme_bw()
```

# Disclaimer

*This software package is developed and maintained by scientists at the NOAA Fisheries Pacific Islands Fisheries Science Center and should be considered a fundamental research communication. The recommendations and conclusions presented here are those of the authors and this software should not be construed as official communication by NMFS, NOAA, or the U.S. Dept. of Commerce. In addition, reference to trade names does not imply endorsement by the National Marine Fisheries Service, NOAA. While the best efforts have been made to insure the highest quality, tools such as this are under constant development and are subject to change.*
